\subsection{Problems}
Less than half of the problems here require the use of the theory developed in
this chapter, which is testament to how rare it is that this theory shows up in
contest.
\begin{result}[Fermat's Christmas Theorem]{\label{r:a:n:pr:1}}
    There are positive integers $a$ and $b$ such that $a^2+b^2=n$ iff all primes
    which are $3\pmod 4$ have even exponents in the prime factorisation of $n$.
    If $n$ is prime, then such a representation is unique.
\end{result}
\begin{itemize}
  \item Let $p$ and $q$ be primes. Prove that there is an integer $x$ such that
    $(x+1)^p\equiv x^p\pmod q$ if and only if $q\equiv 1\pmod p$.
  \item Find all pairs of positive integers $x,y$ such that
    $1+2^x+2^{2x}=y^2$.
  \item Let $p$ be an odd prime. Prove that
    \[p^2\mid 1^p+2^p+\cdots+p^p.\]
  \item Four positive integers $x,y,z,t$ satisfy $xy-zt=x+y=z+t$. Is it possible
    that $xy$ and $zt$ are both perfect squares?
  \item Let $P(x)$ and $Q(x)$ be polynomials whose coefficients are all equal to
    $1$ or $7$. If $P(x)$ divides $Q(x)$, prove that $1+\deg P(x)$ divides
    $1+\deg Q(x)$.
  \item Let $a$ and $b$ be positive integers such that $a^n+n\mid b^n+n$ for all
    positive integers $n$. Prove that $a=b$.
  \item Assume that $g$ is a generator mod $p$ such that $p\mid g^2-g-1$. Prove
      that if $p\equiv 3\pmod 4$, then $g-1$ and $g-2$ are also generators mod
      $p$.
  \item Prove that for all positive integers $a>1$ and $n$ we have
    $n\mid\varphi(a^n-1)$.
\item Determine all positive integers $M$ such that the sequence $a_0, a_1,
      a_2, \ldots$ defined by \[ a_0 = M + \frac{1}{2} \qquad \textrm{and}
        \qquad a_{k+1} = a_k\lfloor a_k \rfloor \quad \textrm{for} \, k = 0, 1,
      2, \ldots \]contains at least one integer term.
  \item Let $r$ be an irrational root of a polynomial $P(x)$ of degree $d$ with
    integer coefficients. Prove that there is a real number $C$ such that for
    any integer $q$ we have $\{qr\}\ge \frac C{q^{d-1}}$.
  \item Find all positive integers $k$ for which the following statement is
    true: if $p$ is an integer polynomial such that $0\le p(i)\le k$ for each
    integer $0\le i\le k+1$, then all of these $p(i)$s are equal.
  \item Let $n$ and $z$ be integers greater than $1$ such that $\gcd(n,z)=1$.
    Prove that there is some nonnegative integer $i<n$ such that
    $1+z+z^2+\cdots+z^i$ is divisible by $n$.
  \item Find the minimum possible value of $m+n$, where $m$ and $n$ are distinct
    positive integers such that $1000\mid 1978^m-1978^n$.
  \item Prove that for some constant $C>0$, the following statement holds:

    Let $m\ge 2$ be an integer, $A$ a finite set of integers (not
    necessarily positive), and $B_1,B_2,\ldots,B_m$ subsets of $A$. Suppose
    that for every $k=1,2,\ldots,m$, the sum of the elements of $B_k$ is
    $2^k$. Then $A$ contains at least $\frac {Cm}{\log_2 m}$ elements.
  \item Find all functions $f:\Zz\to\Zz$ such that, for all integers $a,b,c$
    satsfying $a+b+c=0$, we have
    \[f(a)^2+f(b)^2+f(c)^2=2f(a)f(b)+2f(a)f(c)+2f(b)f(c).\]
  \item Let $p$ be an integer polynomial and let $a$ be an integer such that
      \[
          p(p(\cdots(p(a))\cdots))=a.
      \]
      Prove that $p(p(a))=a$.
  \item Does there exist an integer $n>1$ such that all powers of $n$ are
    base-10 palindromes?
    \item Given three distinct natural numbers $a,b,c$, show that
      \[\gcd(ab+1,bc+1,ca+1)\le\frac{a+b+c}3.\]
  \item Find all integers $x,y$ such that $(x^2+y)(x+y^2)=(x-y)^3$.
  \item Prove that if $5\nmid a$, then $x^5-x+a$ is irreducible.
    \item Find all integers $n$ such that $n^2+1$ divides $n^3+n^2-n-15$.
  \item Find all functions $f:\Nn\to\Nn$ such that $f(f(m)f(n))=mn$ for all
    positive integers $m$ and $n$.
  \item Find all pairs $x,y$ of positive integers such that $3^x-8^y=2xy+1$.
  \item An infinite arithmetic progression contains a perfect $a$th power and a
    perfect $b$th power. Prove that it contains a perfect $\lcm(a,b)$th power.
  \item Let $p$ be an odd prime and let $m$ and $n$ be natural numbers not
    divisible by $p$. Prove that if there is some integer $s$ such that $p\mid
    m^{2^s}+n^{2^s}$, then $p\equiv 1\pmod{2^{s+1}}$.
  \item Let $P(n)$ be the product of the digits of a positive integer $n$.
    Let $n_1$ be a positive integer, and define $n_{i+1}=n_i+P(n_i)$ for each
    $i\ge 1$. Prove that this sequence is eventually constant.
  \item Let $p$ be an odd prime and $r$ an odd natural number. Show that $pr+1$
    does not divide $p^p-1$.
  \item For which positive integers $r$ and $s$ does there exist a positive
    integer $n$ such that $nr$ and $ns$ have the same number of divisors?
  \item Let $a$ and $b$ be positive integers. Prove that there are infinitely
    many positive integers $n$ such that $n^b-1\nmid a^n+1$.
  \item Find all integer polynomials $p$ such that $n\mid p(2^n)$ for all
    positive integers $n$.
  \item Find all pairs of positive integers $x,y$ such that $x^3-y^3=xy+61$.
  \item Find all functions $f:\mathbb N\to\mathbb N$ such that for each positive
    integer $n$, $f(n)\mid n^3$ and $\displaystyle\sum_{i=1}^n f(i)$ is a perfect
    square.
  \item Does there exist a strictly increasing function $f:\Nn\to\Nn$ such that
    $f(1)=2$ and $f(f(n))=f(n)+n$ for all positive integers $n$?
  \item Prove that if there are two terms of an arithmetic progression which are
    coprime integers, then there is an infinite subset of that
    progression all of whose elements are coprime integers.
    \item We call a 5-tuple of integers \emph{arrangeable} if its elements can be
      labelled $a,b,c,d,e$ in some order such that $a-b+c-d+e=0$. Determine all
      2022-tuples of integers such that if we place them in order around a
      circle, then any 5-tuple of numbers in consecutive positions is
      arrangeable.
  \item Let $a,b,c$ be pairwise coprime positive integers. Prove that there
    exist infinitely many triples $x,y,z$ of distinct positive integers such
    that $x^a+y^b=z^c$.
    \item Let $p$ be a prime with $p>3$. Prove that there are positive integers
      $a<b<\sqrt p$ such that $p-b^2\mid p-a^2$.
    \item
        Do there exist distinct prime numbers $a,b,c$ such that
          \[a\mid bc+b+c,\ b\mid ac+a+c,\ c\mid ab+a+b?\]
  \item Let $x,y,z$ be rational numbers such that $xy,\ yz,\ zx,\ x+y+z$ are all
    integers.

    Prove that $x,\ y,\ z$ are all integers.
  \item Find all primes $p,q,r$ such that $p\mid q^r+1,\ q\mid r^p+1,\ r\mid
    p^q-1$.
    \item Let $a$ and $b$ be positive integers. Show that if $4ab-1$ divides
      $(4a^2-1)^2$, then $a=b$.
  \item Let $p$ be a polynomial of dgree $d$. Find a linear
    equation that the values $p(0),p(1),\ldots,p(d+1)$ always satisfy.
  \item Let $k$ and $n$ be positive integers such that $n$ is odd. Prove that
    there is an integer $a$ such that $a^{32}\equiv(n+1)^3\pmod{n^k}$.
  \item Find all functions $f:\Zz_{\ge 0}\to \Zz_{\ge 0}$ such that
    $f(f(f(n)))=f(n+1)+1$ for all $n\in\Zz_{\ge 0}$.
  \item Find all functions $f:\mathbb N\to\mathbb N$ satisfying 
    \[f(n)+f(n+1)=f(n+2)f(n+3)-2023\]
    for all $n\in\mathbb N$.
    \item Prove that for any nonnegative integer $n$, the number $7^{7^n}+1$ is
      the product of at least $2n+3$ (not necessarily distinct) primes.
\item Find all triples $a,b,c$ of positive integers such that
      $a^3+b^3+c^3=(abc)^2$.
  \item Find all positive integers $n$ such that $n\mid 2^{n-1}+1$.
  \item Find all positive integer solutions to $3^x+4^y=5^z$.
    \item Let $n$ and $k$ be positive integers. Assume that for each positive
      integer $m$, there exists a positive integer $a$ such that $a^k\equiv
      n\pmod m$. Prove that $n$ is a perfect $k$th power.
  \item Let $a$ and $b$ be two positive integers. Prove that
    \[a^2+\left\lceil\frac{4a^2}b\right\rceil\] is not a square.
  \item Prove that there exist $2023$ distinct positive integers such that each of them
    divides the sum of the rest.
  \item Let $s(n)$ be the sum of the digits of $n$. Prove that for each positive
    integer $k$ there exists a positive integer $n$ such that $n+s(n)$ equals
    either $k$ or $k+1$.
  \item Find all positive integers $n$ that satisfy the following property:
    for all positive integers $m$, relatively prime to $n$, we have
    $2n^2$ divides $m^n-1$.
  \item Find all natural numbers $n$ such that $n^2\mid 2^n+1$.
  \item What is the smallest positive integer $n$ for which there exist positive
    integers $x_1,x_2,\ldots,x_n$ such that
    \[x_1^3+x_2^3+\cdots+x_n^3=2002^{2002}?\]
  \item Prove that there are infinitely many pairs $a$ and $b$ of perfect
    squares such that they have the same number of digits in decimal, and their
    concatenation is also a square.
  \item Let $n>1$ be an odd positive integer and let $S$ be the set of integers
    $x$, with $1\le x\le n$, such that both $x$ and $x+1$ are coprime to $n$.
    Find the product of the elements of $S$ mod $n$.
  \item Is there a sequence $a_1,\ldots$ of primes such that for each $i$ we
    have $10 a_i\le a_{i+1}<10a_i+9$?
  \item Let $a_1,a_2,a_3,\ldots$ be a sequence of integers, such that for any
    positive integers $n$ and $k$, the quantity
    \[\frac{a_n+a_{n+1}+\cdots+a_{n+k-1}}k\]
    is always the square of an integer. Prove that all $a_i$s are equal.
  \item Let $p>3$ be a prime and let $a,b,c$ be integers with $a\ne 0$. Suppose
    that $ax^2+bx+c$ is a perfect square for $p$ consecutive integers $x$. Prove
    that $p\mid b^2-4ac$.
  \item Prove that all terms of the sequence $a_1=a_2=a_3=1,\
    a_{n+1}=(1+a_{n-1}a_n)/a_{n+2}$ are integers.
    \item Let $a$ be a positive integer such that $4(a^n+1)$ is a perfect
      cube for all positive integers $n$. Prove that $a=1$.
  \item Let $n>1$ be a positive integer and let $p$ be a prime. Given that
    $n\mid p-1$ and $p\mid n^3-1$, prove that $4p-3$ is a perfect square.
  \item Define
    \[q(n)=\left\lfloor\frac n{\lfloor\sqrt n\rfloor}\right\rfloor.\]
    Determine all positive integers $n$ for which $q(n)>q(n+1)$.
    \item Find all pairs of positive integers $a,b$ such that \[b^2-a\mid
        a^2+b\quad\text{and}\quad a^2-b\mid b^2+a.\]
  \item Is there a function $f:\Nn\to\Nn$ of positive integers such that
    $\gcd(a_m,a_n)=1\iff |m-n|=1$?
  \item Show that for any fixed integers $n$ and $a$, the sequence
    $a,a^a,a^{a^a},\ldots$ is eventually constant mod $n$.
  \item Prove that there are infinitely many pairs of positive integers $a,b$
    such that $a\mid b^2+1$ and $b\mid a^2+1$.
  \item Let $k=2^{2^n}+1$ for some positive integer $n$. Prove that $k$ is prime
    if and only if $k\mid 3^{(k-1)/2}+1$.
  \item Let $P$ be a polynomial of degree larger than $1$ with integer
    coefficients. Prove that there are infinitely many positive integers which cannot be
    written in the form $P(x+1)+P(x+2)+\cdots+P(x+k)$ for positive integers $x$
    and $k$.
  \item Let $b\ge 5$ be a fixed positive integer. A positive integer is called a
    base-$b$ palindrome if it reads the same forwards as backwards when written
    in base $b$. 
    
    Prove that there are infinitely many pairs $(x,y)$ of distinct
    perfect squares such that for any base-$b$ palindrome $p$, we have $(x-p)(y-p)>0$.
  \item Prove that if a sequence $\{a_n\}_0^\infty$ of nonnegative integers satisfies
    $a_0=0$ and $a_n=n-a_{a_n}$ for all $n$, then $a_{n+2}>a_n$ for all $n$.
    \item Let $S$ be the set of ordered pairs of integers. We say that two
      elements $(a,b)$ and $(c,d)$ of $S$ are $k$-friends if there is an element
      $(e,f)$ of $S$ such that the area of the triangle formed by these three
      points is $k$.
      Find the smallest positive integer $k$ such that there exists a set of
      $200$ elements of $S$ such that any pair of them are $k$-friends.
  \item Given an integer $k\geq 2$, determine all functions $f:\Nn\to\Nn$
    such that $f(x_1)!+f(x_2)!+\cdots f(x_k)!$
    is divisible by $x_1!+x_2!+\cdots x_k!$ for all positive integers
    $x_1,x_2,\cdots x_k$.
  \item Find all functions $f:\Nn\to\Nn$ such that $f(n+1)>f(f(n))$ for all
    positive integers $n$.
  \item Find all positive integers $k$ such that for all positive integers $n$,
    there exist a prime $p$ and positive integers $x$ and $y$ for which
    $\gcd(x,y)=1$ and $p^n\mid \frac{x^k-y^k}{x-y}$.
  \item Find all functions $f:\Nn\to\Nn$ such that $m^2+f(n)\mid
    mf(m)+n$ for all positive integers $m$ and $n$.
  \item Let $x$ be an irrational number. Prove that there are infinitely many
    positive integers $n$ such that $\{nx\}<\frac 1n$.
  \item Let $p$ be a prime number. Prove that there exists a prime number $q$
    such that for every integer $n$, the number $n^p-p$ is not divisible by $q$.
  \item Let $p\ge 7$ be prime. Prove that for any integer $n$ there are integers $a$ and
    $b$, both not divisible by $p$, such that $p\mid a^2-b^2-n$.
  \item Given a positive integer $k$, show that there exists a prime $p$ such
    that one can choose distinct integers
    $a_1,a_2,\ldots,a_{k+3}\in\{1,2,\ldots,p-1\}$ such that $p$ divides
    $a_i a_{i+1}a_{i+2}a_{i+3}-i$ for all $i=1,2,\ldots,k$.
  \item Find all completely multiplicative functions $f:\mathbb N\to\mathbb Z$
    such that for all $a,b\in\mathbb N$, at least two of $f(a),f(b),f(a+b)$ are
    equal.
  \item Let $n$ be a positive integer, and let $S$ be a set of $n$ positive
    integers all at most $n^2$. Prove that there is a set $T$ of $n$ positive
    integers such that the set $\{s+t:s\in S,t\in T\}$ covers at least half of
    the residues mod $n^2$.
  \item Let $a,b,c$ be integers such that $a/b+b/c+c/a$ and $a/c+c/b+b/a$ are
    both integers. Prove that $|a|=|b|=|c|$.
  \item Let $a,b,n$ be positive integers such that $a>b>1$ and $b$ is odd. If
    $b^n\mid a^n-1$, prove that $na^b>3^n$.
  \item Let $x$ and $y$ be positive integers and let $p$ be prime. Assume there
    are coprime positive integers $m$ and $n$ such that $x^m\equiv y^n\pmod p$.
    Prove that there is a unique positive integer $z$ with $0\le z<p$ such that
    \[x\equiv z^n\pmod p,\qquad y\equiv z^m\pmod p.\]
  \item Find all positive integers $x,y,n$ such that $\gcd(x,n+1)=1$ and
    $x^n+1=y^{n+1}$.
    \item Prove that every positive integer can be uniquely represented as a sum
      of one or more Fibonacci numbers such that the sum does not include two
      consecutive Fibonacci numbers.
  \item Let $n$ and $k$ be positive integers such that $\varphi^k(n)=1$ (that
    is, $\varphi$ iterated $k$ times). Prove that
    $n\le 3^k$.
  \item Find all pairs of positive integers $x,y$ such that
    $1+2^x+2^{2x+1}=y^2$.
  \item Find all pairs of positive integers $x,p$ such that $p$ is prime, $x\le
    2p$, and $x^{p-1}\mid (p-1)^x+1$.
  \item Find all pairs of positive integers $x,y$ such that $xy^2+y+7\mid
    x^2y+x+y$.
  \item Let $p$ be an irreducible integer polynomial. Prove that $p$ does not
    have multiple roots.
  \item Prove that if $a$ and $b$ are positive integers, then $4ab-a-b$ is not a
    perfect square.
\end{itemize}
